---
title: "Statistical assignment 3"
author: "Alicia Rey-Herme (014989)"
date: "21 February 2019"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r}
    
# Install tidyverse and data.table packages
library(tidyverse)
library(data.table)

# Load in data and select variables (dplyr::)
Data8 <- fread("C:/Users/alici/Documents/DA3/data/SN6614/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        dplyr::select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- fread("C:/Users/alici/Documents/DA3/data/SN6614/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        dplyr::select(pidp, sex_dv, ukborn, plbornc)

# Join data
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)

# Recode data
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East (England)",
                         `2` = "North West (England)",
                         `3` = "Yorkshire and The Humber",
                         `4` = "East Midlands (England)",
                         `5` = "West Midlands (England)",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East (England)",
                         `9` = "South West (England)",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )

    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.

(Note the position of the code chunks; each is preceded by four spaces. This helps display numbered lists correctly in the Markdown file with the output.)

1. Histogram (20 points)

    ```{r}
    
Data %>%
  ggplot(aes(x = h_age_dv)) +
  geom_histogram(bins = 87) +
  xlab("Age") +
  ylab("Number of respondents")
         
    ```

2. Scatter plot (20 points). The red line shows a linear fit; the blue line shows a quadratic fit. Note the size and position of points.

    ```{r}

Data %>%
  ggplot(aes(x = h_age_dv, y = h_payn_dv)) +
  geom_point(position = position_jitter(width = 1, height = 1), size = 0.5) +
  xlim(NA, 65) +
  geom_smooth() +
  geom_smooth(method='lm', formula=y~x, col = 'red') +
  xlab("Age") +
  ylab("Monthly earnings")

    ```

3. Faceted density chart (20 points).

    ```{r}
    
Data %>%
  filter(!is.na(placeBorn)) %>%
  ggplot(aes(x = h_age_dv)) +
  geom_density(fill = "black") +
  xlab("Age") +
  ylab("Density") +
  facet_wrap(~ placeBorn, ncol = 3)


    ```

4. Ordered bar chart of summary statistics (20 points).

    ```{r}

Data %>%
  filter(!is.na(sex_dv) & !is.na(placeBorn)) %>%
  group_by(sex_dv, placeBorn) %>%
  summarise(medianEarnings = median(h_payn_dv, na.rm = TRUE)) %>%
  ggplot(aes(x = placeBorn, y = medianEarnings, fill = sex_dv, reorde)) +
  geom_bar(position = "dodge", stat="identity") +
  scale_fill_manual(values = c("dark red","dark blue")) +
  theme(legend.position = "top", legend.title = element_blank()) +
  xlab("Median net monthly earnings") +
  ylab("Country of birth") +
  coord_flip()
    
    Data$
    
Data$placeBorn <- factor(Data$placeBorn, levels = c("Bangladesh", "Pakistan", "Poland", "Jamaica", "India", "UK", "other", "Nigeria", "Ireland"))

  
    ```

5. Map (20 points). This is the most difficult problem in this set. You will need to use the NUTS Level 1 shape file (available here -- https://data.gov.uk/dataset/2aa6727d-c5f0-462a-a367-904c750bbb34/nuts-level-1-january-2018-full-clipped-boundaries-in-the-united-kingdom) and a number of packages for producing maps from shape files. You will need to google additional information; there are multiple webpages with the code that produces similar maps.

    ```{r}
    
# Install necessary packages
library(rgdal) # Load in shapefile
library(maps)

install.packages("leaflet")
library(leaflet)
library(tigris)
    
# Read in the shape file.
area <- readOGR(dsn = "C:/Users/alici/Documents/DA3/assignment4/template", layer = "NUTS_Level_1_January_2018_Full_Clipped_Boundaries_in_the_United_Kingdom")
area <- fortify(area)

# Calculatre median earnings by region 
MapData <- Data %>%
filter(!is.na(h_gor_dv)) %>%
  group_by(h_gor_dv) %>%
  summarise(MedEarningsRegion = median(h_payn_dv, na.rm = TRUE))
MapData

View(Data$h_gor_dv)

MapData2 <- merge(x = area, y=  MapData, by.x = "nuts118nm", by.y = "h_gor_dv")
head(MapData2)

# Plot map
map_base <- MapData2 %>%
  ggplot(mapping = aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "grey") +
  theme(axis.line=element_blank(),
            axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),
          panel.border=element_blank(),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          plot.background=element_blank()) +
  ggtitle("Median earnings by region (£)")
map_base

######## UP TO HERE WORKS

map <- map_base +
      geom_polygon(data = MapData2, aes(fill = MedEarningsRegion, color = NA)) +
      geom_polygon(color = "black", fill = NA) +
      theme_bw() +
map 

  ggplot() +
  geom_polygon(data = area, aes(x = long, y = lat, group = group), fill = Data) +
  scale_colour_brewer(palette="Blues") + 
  title("Median earnings by region (£)") +
  coord_fixed()
map

rm(list=ls())

    ```